GPU-friendly Stroke Expansion

### 2024-04-08 paper.tex
Address review feedback in paper
This is a fairly small set of changes in response to review, including new insight into the hybrid nature of flattening in the Skia and Rive renderers.

We propose that the correctness of stroke outlines be divided into \emph{weak correctness} and \emph{strong correctness.} We define strong correctness as the computation of the outline of a line swept along the segment, maintaining normal orientation, combined with stroke caps and joins. Weak correctness, by contrast, only requires the parallel curves of stroke segments, combined with caps and the outer contours of joins. The two notions are equivalent for sufficiently well-behaved input, in particular when the curvature measured at endpoints of path segments does not exceed the reciprocal of the half linewidth. As described in detail in \cite{Nehab2020}, very few existing implementations actually implement the strong version, so document authors have become accustomed to not depending on behavior at endpoints. Standards for graphics formats also provide an ``out,'' enabling the weaker behavior. [TODO: find language in SVG standard]. While our current implementation emphasizes weak correctness, we believe it can be extended to the stronger sense, by implementing evolutes and inner join contours.

@@

The core problem in stroke expansion is approximating the desired curve by segments of some other curve, usually a simpler one. These segments must be within an error tolerance of the source curve, and ideally there are not many more of them than optimal. We consider a number of source to target pairs, most importantly cubic Béziers to Euler spirals, and Euler spiral parallel curves to either lines or arcs.

There are generally three approaches to such curve approximation. The most straightforward but also least efficient is ``cut then measure,'' usually combined with adaptive subdivision. In this technique, a candidate approximate curve is produced, then the error is measured against the source curve, usually by sampling a number of points along both curves and determining a maximum error (or perhaps some other error norm). If the error is within tolerance, the approximation is accepted. Otherwise, the curve is subdivided (usually at t = 0.5) and each half is recursively approximated. A substantial fraction of all curve approximation methods in the literature are of this form, including \cite{Nehab2020}. The main disadvantage is the cost of computing the error metric. Another risk is underestimating the error due to inadequate sampling; this is a particular problem when the source curve contains a cusp.

@@

By far the most efficient approach is an \emph{invertible} error metric. In this approach, the error metric has an analytic inverse, or at least a good numerical approximation. Because the metric is invertible, it can predict the number of subdivisions needed, as well as the parameter value for each subdivision. If the error metric is accurate, then approximation is near-optimal. One example of an invertible error metric is angle step, used in polar stroking\cite{Kilgard2020}; the number of subdivisions is the total angle subsumed by the curve divided by the angle step size, and the parameter value for each subdivision is the result of solving for a tangent direction. Another widely used invertible error metric is Wang's formula (\cite{Goldman2003}, Section 5.6.3), which gives a bound on the flattening error based on the second derivative of the curve. This metric is conservative but works well in practice; among other applications, it is used in Skia for path flattening. One limitation of Wang's method is that it only applies to the flattening of filled outlines, not parallel curves as is required for strokes.

@@

Using a subpixel Fréchet distance bound guarantees that the rendered image does not deviate visibly from the exact rendering. Another choice would be uniform steps in tangent angle, as chosen by \cite{Kilgard2020}. However, at small curvature, the stroked path can be off by several pixels, and at large curvature there may be considerably more subdivision than needed for faithful rendering.

>>>>

We propose that the correctness of stroke outlines be divided into \emph{weak correctness} and \emph{strong correctness.} We define strong correctness as the computation of the outline of a line swept along the segment, maintaining normal orientation, combined with stroke caps and joins. Weak correctness, by contrast, only requires the parallel curves of stroke segments, combined with caps and the outer contours of joins. The two notions are equivalent for sufficiently well-behaved input, in particular when the curvature measured at endpoints of path segments does not exceed the reciprocal of the half linewidth. As described in detail in \cite{Nehab2020}, very few existing implementations actually implement the strong version, so document authors have become accustomed to not depending on behavior at endpoints. Standards for graphics formats, such as SVG\cite{Svg2}, also provide an ``out,'' enabling the weaker behavior. While our current implementation emphasizes weak correctness, we believe it can be extended to the stronger sense, by implementing evolutes and inner join contours.

@@

The core problem in stroke expansion is approximating the desired curve by segments of some other curve, usually a simpler one. These segments must be within an error tolerance of the source curve, and ideally close to a minimal number of them. We consider a number of source to target pairs, most importantly cubic Béziers to Euler spirals, and Euler spiral parallel curves to either lines or arcs.

There are generally three approaches to such curve approximation. The most straightforward but also least efficient is ``cut then measure,'' usually combined with adaptive subdivision. In this technique, a candidate approximate curve is produced, then the error is measured against the source curve, usually by sampling a number of points along both curves and determining a maximum error (or perhaps some other error norm). If the error is within tolerance, the approximation is accepted. Otherwise, the curve is subdivided (usually at $t = 0.5$) and each half is recursively approximated. A substantial fraction of all curve approximation methods in the literature are of this form, including \cite{Nehab2020}. The main disadvantage is the cost of computing the error metric. Another risk is underestimating the error due to inadequate sampling; this is a particular problem when the source curve contains a cusp.

@@

By far the most efficient approach is an \emph{invertible} error metric. In this approach, the error metric has an analytic inverse, or at least a good numerical approximation. Because the metric is invertible, it can predict the number of subdivisions needed, as well as the parameter value for each subdivision. If the error metric is accurate, then approximation is near-optimal. One example of an invertible error metric is angle step, used in polar stroking\cite{Kilgard2020}; the number of subdivisions is the total angle subsumed by the curve divided by the angle step size, and the parameter value for each subdivision is the result of solving for a tangent direction. Another widely used invertible error metric is Wang's formula (\cite{Goldman2003}, Section 5.6.3), which gives a bound on the flattening error based on the second derivative of the curve. This metric is conservative but works well in practice; among other applications, it is used in Skia for path flattening. One limitation of Wang's method is that it only applies to the flattening of filled outlines; when applied naively to generation of parallel curves, it can undershoot substantially, especially near cusps.

@@

Using a subpixel Fréchet distance bound guarantees that the rendered image does not deviate visibly from the exact rendering. Another choice would be uniform steps in tangent angle, as chosen by polar stroking\cite{Kilgard2020}. However, at small curvature, the stroked path can be off by several pixels, and at large curvature there may be considerably more subdivision than needed for faithful rendering. It should be noted, to avoid the undershoot at low curvature, both the Skia\cite{Skia} and Rive\cite{Rive} renderers use a hybrid of the Wang and polar stroking error metrics.
